---
title: "CHAT Trial Primary Analysis - CONSORT Participant Flow"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---

# Summary

Track participant flow in the CHAT trial.

Summarize patient numbers by treatment group and phase to include in a CONSORT flow diagram.


# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R","CHAT-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHAT-primary-data-processing.R`.

```{r load the data, warning= FALSE, message = FALSE}
#---------------------------------
# load the preprocessed data
#---------------------------------
# precensus data
precensus <- read_rds(here("data","CHAT_precensus_public.rds"))

dch <- read_rds(here("data","CHAT_child_census_public.rds"))

dch_wide <- read_rds(here("data","CHAT_child_census_wide_public.rds"))

# child-phase level data
dchp <- read_rds(here("data","CHAT_child_phase_public.rds"))
  
# cluster-phase level summaries
dclp <- read_rds(here("data","CHAT_cluster_phase_public.rds"))

# cluster-level summaries
dcl <- read_rds(here("data","CHAT_cluster_public.rds"))

# Treatment allocation
tx_alloc <- read_rds(here("data","CHAT_tx_alloc_public.rds"))
```

```{r process the data, warning= FALSE, message = FALSE}
#------------------------------
# convert person-days into
# person-years for convenience
#------------------------------
dchp2 <- dchp %>%
  mutate(person_years = person_days/365.25)

dclp2 <- dclp %>%
  mutate(person_years = person_days/365.25)

dcl2 <- dcl %>%
  mutate(person_years = person_days/365.25)

#------------------------------
# for the cluster-level data
# drop clusters with no person
# time for the efficacy and
# futility analyses
#------------------------------
dclp3 <- dclp2 %>%
  filter(person_years > 0)

dcl3 <- dcl2 %>%
  filter(person_years > 0)

```

```{r, warning= FALSE, message = FALSE}
dchp_long <- dch_wide %>%
  select(clusterid_public,tx, childid_public, contains("vital"), contains("agem")) %>%
  distinct() %>%
  left_join(
    dcl2 %>% select(clusterid_public, contains("cl_meas")) %>%
    mutate_if(is.numeric, ~replace_na(., 0))
  ) %>%
  mutate(
    vital6 = ifelse(vital0 == "alive" & agem0 %in% 1:59 & cl_meas6 == 1 & is.na(vital6), "missing", vital6),
    vital12 = ifelse(vital6 == "alive" & agem6 %in% 1:59 &cl_meas12 == 1 & is.na(vital12), "missing", vital12),
    vital18 = ifelse(vital12 == "alive" & agem12 %in% 1:59 &cl_meas18 == 1 & is.na(vital18), "missing", vital18),
    vital24 = ifelse(vital18 == "alive" & agem18 %in% 1:59 &cl_meas24 == 1 & is.na(vital24), "missing", vital24),
    vital30 = ifelse(vital24 == "alive" & agem24 %in% 1:59 &cl_meas30 == 1 & is.na(vital30), "missing", vital30),
    vital36 = ifelse(vital30 == "alive" & agem30 %in% 1:59 &cl_meas36 == 1 & is.na(vital36), "missing", vital36),
    vital6 = ifelse(vital0 == "alive" & agem0 %in% 1:59 & cl_meas6 == 0 & is.na(vital6), "cl_notcensused", vital6),
    vital12 = ifelse(vital6 == "alive" & agem6 %in% 1:59 & cl_meas12 == 0 & is.na(vital12), "cl_notcensused", vital12),
    vital18 = ifelse(vital12 == "alive" & agem12 %in% 1:59 & cl_meas18 == 0 & is.na(vital18), "cl_notcensused", vital18),
    vital24 = ifelse(vital18 == "alive" & agem18 %in% 1:59 & cl_meas24 == 0 & is.na(vital24), "cl_notcensused", vital24),
    vital30 = ifelse(vital24 == "alive" & agem24 %in% 1:59 & cl_meas30 == 0 & is.na(vital30), "cl_notcensused", vital30),
    vital36 = ifelse(vital30 == "alive" & agem30 %in% 1:59 & cl_meas36 == 0 & is.na(vital36), "cl_notcensused", vital36)
  ) %>%
  select(-contains("cl_meas")) %>%
  pivot_longer(
    cols = starts_with("vital"),
    names_to = "timepoint",
    values_to = "vital",
    names_prefix = "vital"
  ) %>% 
  pivot_longer(
    cols = starts_with("agem"),
    names_to = "timepoint_agem",
    values_to = "agem",
    names_prefix = "agem"
  ) %>%
  filter(timepoint == timepoint_agem)%>%
  select(-timepoint_agem)
```

# CONSORT flow

Summarize the number of clusters randomized, and the number that contributed to each study phase. The total number of clusters contributing to each phase is lower than the number randomized because of instability in Burkina Faso that prevented field teams from safely visiting a portion of the study region.

## Census Round

```{r, warning= FALSE, message = FALSE}

# randomization dataset
consort_tx <- tx_alloc %>% 
  group_by(tx) %>% 
  summarize(nrandomized = n()) %>% 
  adorn_totals("row")

# precensus
consort_precensus <- precensus %>% 
  group_by(tx) %>% 
  summarize(n_precensused_cl = n_distinct(clusterid_public),
            n_precensused_ch = sum(childUnder5)) %>% 
  adorn_totals("row")

# clusters in each round of census
consort_census_cl <- tx_alloc %>%
  left_join(dcl2 %>% select(- tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  group_by(tx) %>% 
  select(tx,starts_with("cl_meas")) %>% 
  summarize_all(function(x) {
    sum(x)
  }) %>%
  adorn_totals("row") %>%
  rename_with(~ gsub("cl_meas", "cl_censused_", .x), everything())

# clusters lost to follow up in each round of census
consort_census_clltfu <- tx_alloc %>%
  left_join(dcl2 %>% select(- tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  group_by(tx) %>% 
  select(tx,starts_with("cl_meas")) %>% 
  summarise(
    cl_ltfu_0 = paste0(
      sum(1 - cl_meas0), 
      " (", round(sum(1 - cl_meas0) / n() * 100, 1), "%)"
    ),
    cl_ltfu_6 = paste0(
      sum(cl_meas0 == 1 & cl_meas6 == 0), 
      " (", round(sum(cl_meas0 == 1 & cl_meas6 == 0) / sum(cl_meas0 == 1) * 100, 1), "%)"
    ),
    cl_ltfu_12 = paste0(
      sum(cl_meas6 == 1 & cl_meas12 == 0), 
      " (", round(sum(cl_meas6 == 1 & cl_meas12 == 0) / sum(cl_meas6 == 1) * 100, 1), "%)"
    ),
    cl_ltfu_18 = paste0(
      sum(cl_meas12 == 1 & cl_meas18 == 0), 
      " (", round(sum(cl_meas12 == 1 & cl_meas18 == 0) / sum(cl_meas12 == 1) * 100, 1), "%)"
    ),
    cl_ltfu_24 = paste0(
      sum(cl_meas18 == 1 & cl_meas24 == 0), 
      " (", round(sum(cl_meas18 == 1 & cl_meas24 == 0) / sum(cl_meas18 == 1) * 100, 1), "%)"
    ),
    cl_ltfu_30 = paste0(
      sum(cl_meas24 == 1 & cl_meas30 == 0), 
      " (", round(sum(cl_meas24 == 1 & cl_meas30 == 0) / sum(cl_meas24 == 1) * 100, 1), "%)"
    ),
    cl_ltfu_36 = paste0(
      sum(cl_meas30 == 1 & cl_meas36 == 0), 
      " (", round(sum(cl_meas30 == 1 & cl_meas36 == 0) / sum(cl_meas30 == 1) * 100, 1), "%)"
    )
  ) %>%
  bind_rows(
     tx_alloc %>%
      left_join(dcl2 %>% select( - tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
      mutate_if(is.numeric, ~replace_na(., 0)) %>%
      select(tx,starts_with("cl_meas")) %>% 
      summarise(
        cl_ltfu_0 = paste0(
          sum(1 - cl_meas0), 
          " (", round(sum(1 - cl_meas0) / n() * 100, 1), "%)"
        ),
        cl_ltfu_6 = paste0(
          sum(cl_meas0 == 1 & cl_meas6 == 0), 
          " (", round(sum(cl_meas0 == 1 & cl_meas6 == 0) / sum(cl_meas0 == 1) * 100, 1), "%)"
        ),
        cl_ltfu_12 = paste0(
          sum(cl_meas6 == 1 & cl_meas12 == 0), 
          " (", round(sum(cl_meas6 == 1 & cl_meas12 == 0) / sum(cl_meas6 == 1) * 100, 1), "%)"
        ),
        cl_ltfu_18 = paste0(
          sum(cl_meas12 == 1 & cl_meas18 == 0), 
          " (", round(sum(cl_meas12 == 1 & cl_meas18 == 0) / sum(cl_meas12 == 1) * 100, 1), "%)"
        ),
        cl_ltfu_24 = paste0(
          sum(cl_meas18 == 1 & cl_meas24 == 0), 
          " (", round(sum(cl_meas18 == 1 & cl_meas24 == 0) / sum(cl_meas18 == 1) * 100, 1), "%)"
        ),
        cl_ltfu_30 = paste0(
          sum(cl_meas24 == 1 & cl_meas30 == 0), 
          " (", round(sum(cl_meas24 == 1 & cl_meas30 == 0) / sum(cl_meas24 == 1) * 100, 1), "%)"
        ),
        cl_ltfu_36 = paste0(
          sum(cl_meas30 == 1 & cl_meas36 == 0), 
          " (", round(sum(cl_meas30 == 1 & cl_meas36 == 0) / sum(cl_meas30 == 1) * 100, 1), "%)"
        )
      )  %>%
      mutate(tx = "Total")
  )

# children in each round of census
consort_census_ch <- dchp_long %>%
  # exclude dead child in baseline - never been in study
  filter(! (vital %in% c("dead", "moved","unknown", "missing", "cl_notcensused") & timepoint == "0"))%>%
  group_by(timepoint, tx) %>%
  summarise(
    ch_censused = sum(vital %in% c("alive","dead") & agem %in% 1:59, na.rm = TRUE),
    ch_moved_unknown = sum(vital %in% c("moved","unknown", "missing"), na.rm = TRUE),
    ch_cl_notcensused = sum(vital %in% c("cl_notcensused"), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(timepoint = factor(gsub("vital", "", timepoint),
                         c("0", "6", "12", "18", "24", "30", "36"))) %>%
  arrange(timepoint) %>%
  pivot_wider(
    names_from = timepoint,
    values_from = c(ch_censused, ch_moved_unknown, ch_cl_notcensused),
    names_glue = "{.value}_{timepoint}"
  ) %>% 
  adorn_totals("row")


consort_census <- consort_tx %>%
  full_join(consort_precensus) %>%
  full_join(consort_census_cl) %>%
  full_join(consort_census_clltfu) %>%
  full_join(consort_census_ch) %>%
  select(
    tx, nrandomized,
    n_precensused_cl, n_precensused_ch, 
    cl_censused_0, cl_ltfu_0, ch_censused_0, ch_moved_unknown_0, ch_cl_notcensused_0, 
    cl_censused_6, cl_ltfu_6, ch_censused_6, ch_moved_unknown_6, ch_cl_notcensused_6, 
    cl_censused_12, cl_ltfu_12, ch_censused_12, ch_moved_unknown_12, ch_cl_notcensused_12, 
    cl_censused_18, cl_ltfu_18, ch_censused_18, ch_moved_unknown_18, ch_cl_notcensused_18, 
    cl_censused_24, cl_ltfu_24, ch_censused_24, ch_moved_unknown_24, ch_cl_notcensused_24, 
    cl_censused_30, cl_ltfu_30, ch_censused_30, ch_moved_unknown_30, ch_cl_notcensused_30, 
    cl_censused_36, cl_ltfu_36, ch_censused_36, ch_moved_unknown_36, ch_cl_notcensused_36
  ) %>%
  # transpose
  t() %>% 
  # convert 1st row (tx) to colnames 
  row_to_names(1) %>%
  # convert matrix output of transpose to df 
  as.data.frame() %>% 
  # rename Total appropriately
  rename(Overall = Total) %>% 
  # convert rownames to columns 
  # necessary because rownames would otherwise need to be unique
  # set column name to empty string
  rownames_to_column(" ") %>% 
  # rename variables for presentation
  mutate(` ` = case_when(` ` == "nrandomized" ~ "No. of clusters",
                         ` ` == "n_precensused_cl" ~ "No. of clusters",
                         ` ` == "n_precensused_ch" ~ "No. of children",
                         str_detect(` `, "cl_censused") ~ "Clusters censused",
                         str_detect(` `, "cl_ltfu") ~ "Clusters lost to follow-up",
                         str_detect(` `, "ch_censused") ~ "Children censused",
                         str_detect(` `, "ch_moved_unknown") ~ "Children moved/unknown",
                         str_detect(` `, "ch_cl_notcensused") ~ "Children lost to follow-up due to cluster lost"
                         )
         ) 

consort_census %>% knitr::kable(format = "html", digits = 0, format.args = list(big.mark = ",")) %>%
            kable_styling(bootstrap_options = "striped") %>%
            pack_rows("Randomization", 1, 1) %>%   
            pack_rows("Precensus", 2, 3) %>%   
            pack_rows("Census 0", 4, 8) %>%
            pack_rows("Census 6", 9, 13) %>% 
            pack_rows("Census 12", 14, 18) %>% 
            pack_rows("Census 18", 19, 23) %>% 
            pack_rows("Census 24", 24, 28) %>% 
            pack_rows("Census 30", 29, 33) %>% 
            pack_rows("Census 36", 34, 38) 
```


## Phase

```{r consort flow Phase, warning= FALSE, message = FALSE}
#------------------------------
# summarize datasets to arrive to consort variables
#------------------------------

# randomization dataset
consort_tx <- tx_alloc %>% 
  group_by(tx) %>% 
  summarize(nrandomized = n()) %>% 
  adorn_totals("row")

# randomization dataset
consort_precensus <- precensus %>% 
  group_by(tx) %>% 
  summarize(n_precensused_cl = n_distinct(clusterid_public),
            n_precensused_ch = sum(childUnder5)) %>% 
  adorn_totals("row")

# phase 0 to 6  
consort_0to6 <- dclp3 %>%
  filter(phase == "0to6") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_0to6 = n_distinct(clusterid_public),
            ch_contribute_0to6 = sum(natrisk),
            pt_years_0to6 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# phase 6 to 12
consort_6to12 <- dclp3 %>%
  filter(phase == "6to12") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_6to12 = n_distinct(clusterid_public),
            ch_contribute_6to12 = sum(natrisk),
            pt_years_6to12 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# phase 12 to 18
consort_12to18 <- dclp3 %>%
  filter(phase == "12to18") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_12to18 = n_distinct(clusterid_public),
            ch_contribute_12to18 = sum(natrisk),
            pt_years_12to18 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# phase 18 to 24
consort_18to24 <- dclp3 %>%
  filter(phase == "18to24") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_18to24 = n_distinct(clusterid_public),
            ch_contribute_18to24 = sum(natrisk),
            pt_years_18to24 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# phase 24 to 30
consort_24to30 <- dclp3 %>%
  filter(phase == "24to30") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_24to30 = n_distinct(clusterid_public),
            ch_contribute_24to30 = sum(natrisk),
            pt_years_24to30 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# phase 30 to 36
consort_30to36 <- dclp3 %>%
  filter(phase == "30to36") %>% 
  group_by(tx) %>% 
  summarise(cl_contribute_30to36 = n_distinct(clusterid_public),
            ch_contribute_30to36 = sum(natrisk),
            pt_years_30to36 = round(sum(person_years), digits = 0)) %>% 
  adorn_totals("row")

# all phases (0 to 12)
consort_0to36 <- dclp3 %>%
  group_by(tx) %>%
  summarise(cl_contribute = n_distinct(clusterid_public),
            pt_years = round(sum(person_years), digits = 0)) %>% 
  inner_join(dchp2 %>% 
    filter(person_days > 0) %>%
    group_by(tx) %>%
    summarise(ch_contribute = n_distinct(childid_public))
    ) %>%
  select(tx, cl_contribute, pt_years, ch_contribute) %>%
  adorn_totals("row")

consort_0to36_age <- dchp2 %>% 
  filter(person_days > 0) %>%
  group_by(childid_public) %>%
  mutate(phase_numeric = case_when(
        phase == "0to6" ~ 1,
        phase == "6to12" ~ 2,
        phase == "12to18" ~ 3,
        phase == "18to24" ~ 4,
        phase == "24to30" ~ 5,
        phase == "30to36" ~ 6
    )) %>%
  filter(phase_numeric == min(phase_numeric)) %>%
  ungroup() %>%
  group_by(tx) %>%
  summarise(ch_age = paste0(round(mean(age_months), 1),
                            " (", round(sd(age_months),1),
                             ")"))%>%
  bind_rows(
    dchp2 %>% 
      filter(person_days > 0) %>%
      group_by(childid_public) %>%
      mutate(phase_numeric = case_when(
            phase == "0to6" ~ 1,
            phase == "6to12" ~ 2,
            phase == "12to18" ~ 3,
            phase == "18to24" ~ 4,
            phase == "24to30" ~ 5,
            phase == "30to36" ~ 6
        )) %>%
      filter(phase_numeric == min(phase_numeric)) %>%
      ungroup() %>%
      summarise(ch_age = paste0(round(mean(age_months), 1),
                                " (", round(sd(age_months),1),
                                 ")"))%>%
      mutate(tx = "Total")
  )

consort_py_0to36 <- dcl3 %>%
  group_by(tx) %>%
  summarise(mean_py = paste0(round(mean(person_years),1), 
                             " (", round(sd(person_years),1),
                             ")"),
            median_py = paste0(round(median(person_years),1),
                            " (", round(quantile(person_years, 0.25),1),
                            ", ", round(quantile(person_years, 0.75),1),
                            ")")
            ) %>%
  bind_rows(
    dcl3 %>%
      summarise(mean_py = paste0(round(mean(person_years),1), 
                             " (", round(sd(person_years),1),
                             ")"),
                median_py = paste0(round(median(person_years),1),
                            " (", round(quantile(person_years, 0.25),1),
                            ", ", round(quantile(person_years, 0.75),1),
                            ")")
            ) %>%
      mutate(tx = "Total")
      
  )

consort_ltfu_0to36 <- tx_alloc %>%
  left_join(dcl3 %>% select(- tx), by = c("clusterid_public"="clusterid_public")) %>%
  left_join(precensus %>% select(- tx), by = "clusterid_public") %>%
  filter(is.na(person_years) | person_years == 0) %>%
  group_by(tx) %>%
  summarise(N_ltfu_cl = n(),
            N_ltfu_ch = sum(childUnder5, na.rm = T)) %>%
  adorn_totals("row")

consort_ltfu_0to36_byphase <- tx_alloc %>%
  left_join(dcl3 %>% select(- tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  group_by(tx) %>% 
  select(tx,starts_with("cl_meas")) %>% 
  summarize_all(function(x) {
    s <- sum(1 - x)
    m <- round(100 * mean(1 - x), 1)
    paste(s, "(", m, "%)", sep = "")
  }) %>%
  bind_rows(
    tx_alloc %>%
  left_join(dcl3 %>% select(- tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  select(starts_with("cl_meas")) %>% 
  summarize_all(function(x) {
    s <- sum(1 - x)
    m <- round(100 * mean(1 - x), 1)
    paste(s, " (", m, "%)", sep = "")
  }) %>%
    mutate(tx = "Total")
  )


consort_nevercensus_0to36 <- tx_alloc %>%
  left_join(dcl2 %>% select(- tx, -csps_distance), by = c("clusterid_public"="clusterid_public")) %>%
  mutate_if(is.numeric, ~replace_na(., 0)) %>%
  group_by(tx) %>% 
  select(tx,starts_with("cl_meas")) %>% 
  summarize(N_ncensus_cl = sum(cl_meas0 == 0 & cl_meas6 == 0 & cl_meas12 == 0 & cl_meas18 == 0 & 
                cl_meas24 == 0 & cl_meas30 == 0 & cl_meas36 == 0 )) %>%
  adorn_totals("row")

#------------------------------
# join all consort tables, by tx
# then manipulate to arrive to final format
#------------------------------
consort <- consort_tx %>% 
  full_join(consort_precensus) %>%
  full_join(consort_0to6) %>% 
  full_join(consort_6to12) %>%
  full_join(consort_12to18) %>%
  full_join(consort_18to24) %>%
  full_join(consort_24to30) %>%
  full_join(consort_30to36) %>%
  full_join(consort_0to36) %>%
  full_join(consort_0to36_age) %>%
  full_join(consort_py_0to36) %>%
  full_join(consort_nevercensus_0to36) %>%
  full_join(consort_ltfu_0to36) %>%
  full_join(consort_ltfu_0to36_byphase) 

consort %<>% 
  # transpose
  t() %>% 
  # convert 1st row (tx) to colnames 
  row_to_names(1) %>%
  # convert matrix output of transpose to df 
  as.data.frame() %>% 
  # rename Total appropriately
  rename(Overall = Total) %>% 
  # convert rownames to columns 
  # necessary because rownames would otherwise need to be unique
  # set column name to empty string
  rownames_to_column(" ") %>% 
  # rename variables for presentation
  mutate(` ` = case_when(` ` == "nrandomized" ~ "No. of clusters",
                         ` ` == "n_precensused_cl" ~ "No. of clusters",
                         ` ` == "n_precensused_ch" ~ "No. of children",
                         str_detect(` `, "cl_contribute") ~ "Clusters contributing",
                         str_detect(` `, "pt_years") ~ "Person years at risk",
                         str_detect(` `, "ch_contribute") ~ "Children contributing",
                         str_detect(` `, "ch_age") ~ "Children's age (SD)",
                         ` ` == "pt_years" ~ "Person-years",
                         ` ` == "mean_py" ~ "Mean person-years per cluster (SD)",
                         ` ` == "median_py" ~ "Median person-years per cluster (Q1, Q3)",
                         ` ` == "N_ltfu_cl" ~ "No. of clusters (not contributing to analysis)",
                         ` ` == "N_ltfu_ch" ~ "No. of children",
                         ` ` == "N_ncensus_cl"~ "No. of clusters never reached",
                         str_detect(` `, "cl_meas") ~ str_replace(` `, "cl_meas", "Clusters lost in CHAT")
                         )
         ) 
# %>%
#   # format cols numeric to help with printing
#   mutate(Placebo = as.numeric(Placebo),
#          Azithromycin = as.numeric(Azithromycin),
#          Overall = as.numeric(Overall))

#------------------------------
# generate table
#------------------------------

consort %>% knitr::kable(format = "html", digits = 0, format.args = list(big.mark = ",")) %>%
            kable_styling(bootstrap_options = "striped") %>%
            pack_rows("Randomization", 1, 1) %>%   
            pack_rows("Precensus", 2, 3) %>%   
            pack_rows("Phase 0-6", 4, 6) %>%
            pack_rows("Phase 6-12", 7, 9) %>% 
            pack_rows("Phase 12-18", 10, 12) %>% 
            pack_rows("Phase 18-24", 13, 15) %>% 
            pack_rows("Phase 24-30", 16, 18) %>% 
            pack_rows("Phase 30-36", 19, 21) %>% 
            pack_rows("Primary analysis, Phases 0-36", 22, 27) %>%
            pack_rows("Primary analysis, Lost to follow-up", 28, 37)

```


# Session Info
```{r session info}
sessionInfo()
```

