---
title: "CHAT Trial Primary Analysis - estimate coefficient of variation"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

Estimate the coefficient of variation in cluster-level mortality rates using an approach proposed by Hayes and Moulton (2017) 7.17: 

$$
\widehat{\sigma}_B^2 = s^2 - \frac{r}{\overline{y}_H}
$$

Where $s^2$ is the between-cluster variance in the mortality rate, $r$ is the overall mortality rate computed from all clusters combined; ${\overline{y}_H}$ is the harmonic mean of the $y_i$ (person-years of follow-up in clusters).

The coefficient of variation is: $$\widehat{k} = \frac{\widehat{\sigma}_B}{r}$$


# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here::here("R","CHAT-primary-Config.R"))
```

# Load the data

Load the final cluster level dataset for the CHAT trial. 

```{r load the data}
#---------------------------------
# load the preprocessed cluster level data
#---------------------------------

# cluster-level dataset
dcl <- read_rds(here("data","CHAT_cluster_public.rds"))

```

```{r process the data}
#------------------------------
# convert person-days into
# person-years for convenience
#
# limit to cluster-phases where children 
# were at risk and were not lost
#
# estimate the cluster-level
# mortality rate per year
#------------------------------
dcl4 <- dcl %>%
  mutate(person_years = person_days/365.25) %>%
  filter(person_years > 0) %>%
  mutate(mrate = ndeaths / person_years)

```

# Estimate coefficient of variation

Estimate it over all clusters, and for each treatment group.

```{r estimate cv all clusters}
cv_all <- dcl4 %>%
  ungroup() %>%
  summarise(`N clusters` = n(),
            `r` = round(sum(ndeaths)/sum(person_years),5),
            `cluster SD(r)` = round(sd(mrate),4),
            `y_H` = round(1/mean(1/person_years), 1),
            `r/y_H` = r/y_H,
            `variance_B` = (sd(mrate))^2 - (r/y_H),
            `CV (k)` = round(sqrt(variance_B)/r,2))

```

```{r estimate cv placebo clusters}
cv_pl <- dcl4 %>%
  ungroup() %>%
  filter(tx == "Placebo") %>%
  summarise(`N clusters` = n(),
            `r` = round(sum(ndeaths)/sum(person_years),5),
            `cluster SD(r)` = round(sd(mrate),4),
            `y_H` = round(1/mean(1/person_years), 1),
            `r/y_H` = r/y_H,
            `variance_B` = (sd(mrate))^2 - (r/y_H),
            `CV (k)` = round(sqrt(variance_B)/r,2))
 
```

```{r estimate cv azithro clusters}
cv_az <- dcl4 %>%
  ungroup() %>%
  filter(tx == "Azithromycin") %>%
  summarise(`N clusters` = n(),
            `r` = round(sum(ndeaths)/sum(person_years),5),
            `cluster SD(r)` = round(sd(mrate),4),
            `y_H` = round(1/mean(1/person_years), 1),
            `r/y_H` = r/y_H,
            `variance_B` = (sd(mrate))^2 - (r/y_H),
            `CV (k)` = round(sqrt(variance_B)/r,2))
```

## Summary table

```{r summary table}
cv_table <- bind_rows(cv_all, cv_pl, cv_az) %>%
  mutate(group = c("All clusters","Placebo","Azithromycin")) %>%
  select(group, everything())

kbl(cv_table) %>%
  kable_styling(bootstrap_options = "striped") 

```

# Session info
```{r session info}
sessionInfo()
```

