---
title: "CHAT Trial Primary Analysis - Mortality"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

This script estimates compares 1-59m mortality between azithromycin and placebo the CHAT trial, the primary endpoint. 

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R","CHAT-primary-Config.R"))
source(here::here("R","CHAT-primary-Functions.R"))
```

# Load the data


```{r load the data}
#---------------------------------
# load the preprocessed data
#---------------------------------

# cluster-level summaries
dcl <- read_rds(here("data","CHAT_cluster_public.rds"))

# treatment allocations
tx_alloc <- read_rds(here("data","CHAT_tx_alloc_public.rds")) %>% 
  dplyr::select(clusterid_public, hdss_strata = mortplus)

# merge on the treatment allocation and strata to the main cluster-level summaries
dcl <- dcl %>%
  left_join(tx_alloc, by="clusterid_public")
  
```

```{r process the data}
#------------------------------
# convert person-days into
# person-years for convenience
#------------------------------
dcl2 <- dcl %>%
  mutate(person_years = person_days/365.25)

#------------------------------
# for cluster-level summaries:
# tabulate and then 
# drop clusters with no person time
#------------------------------

dcl3 <- dcl2 %>%
  filter(person_years > 0)

#------------------------------
# estimate the cluster-level
# mortality rate
#------------------------------
dcl3$mrate <- dcl3$ndeaths/dcl3$person_years
```

# Cluster-level summary distributions

## Pooled distributions

```{r distributions overall}
#-------------------------------
# histogram of person years
# and deaths overall
#-------------------------------
dcl3 %>%
  ggplot(aes(x=person_years))+
  geom_histogram(binwidth=40, alpha=0.2,color="gray40")+
  scale_x_continuous(breaks=seq(0,1200,by=200)) +
  scale_y_continuous(breaks=seq(0,25,by=5)) +
  theme_minimal() +
  theme(legend.position = "none", strip.text.y = element_text(angle=0, size=12, hjust=0)) +
  labs(x="Person years at risk (cluster totals)", y = "Number of clusters",
       title = "Person years at risk", subtitle = "(bin width = 40)")

dcl3 %>%
  ggplot(aes(x=ndeaths))+
  geom_histogram(binwidth=1,alpha=0.2,color="gray40")+
  scale_x_continuous(breaks=seq(0,25,by=5)) +
  scale_y_continuous(breaks=seq(0,55,by=5)) +
  theme_minimal() +
  theme(legend.position = "none", 
        strip.text.y = element_text(angle=0, size=12, hjust=0),
        panel.grid.minor.y = element_blank()) +
  labs(x="Number of deaths", y = "Number of clusters",
       title = "Number of deaths per cluster", subtitle = "(bin width = 1)")

#-------------------------------
# histogram of mortality rates
# overall
#-------------------------------
ggplot(data=dcl3, aes(x = mrate*1000)) +
  geom_histogram(show.legend = TRUE, alpha=0.2, binwidth = 1,color="gray40") +
  scale_x_continuous(breaks = seq(0,50,by=10)) +
  scale_y_continuous(breaks=seq(0,55,by=5)) +
  labs(title="Mortality rate", subtitle = "(histogram bin width is 1)", x="mortality rate (deaths per 1000 child-years)", y="number of clusters") +
  theme_minimal() +
  theme(legend.position = "none", 
        strip.text.y = element_text(angle=0),
        panel.grid.minor.y = element_blank()
        )

```

## Arm-stratified distributions

Person time, Number of Deaths, and Mortality Rates by arm

```{r arm stratified distributions}
#-------------------------------
# histogram of person-years
# by arm
#-------------------------------
dcl3 %>%
  ggplot(aes(x=person_years, color=tx, fill=tx))+
  geom_histogram(binwidth=40, alpha=0.5)+
  scale_color_manual(values=c("#999999", "#E69F00"))+
  scale_fill_manual(values=c("#999999", "#E69F00"))+
  scale_x_continuous(breaks=seq(0,1200,by=200)) +
  facet_grid(tx ~ .) +
  theme_minimal() +
  theme(legend.position = "none", strip.text.y = element_text(angle=0, size=12, hjust=0)) +
  labs(x="Person years at risk (cluster totals)", y = "Number of clusters",
       title = "Person years at risk", subtitle = "(bin width = 40)")

#-------------------------------
# histogram of deaths
# by arm
#-------------------------------
dcl3 %>%
  ggplot(aes(x=ndeaths, color=tx, fill=tx))+
  geom_histogram(binwidth=1,alpha=0.5)+
  scale_color_manual(values=c("#999999", "#E69F00"))+
  scale_fill_manual(values=c("#999999", "#E69F00"))+
  scale_x_continuous(breaks=seq(0,25,by=5)) +
  scale_y_continuous(breaks=seq(0,25,by=5)) +
  facet_grid(tx ~ .) +
  theme_minimal() +
  theme(legend.position = "none", 
        strip.text.y = element_text(angle=0, size=12, hjust=0),
        panel.grid.minor.y = element_blank()) +
  labs(x="Number of deaths (cluster totals)", y = "Number of clusters",
       title = "Number of deaths per cluster", subtitle = "(bin width = 1)")

#-------------------------------
# histogram of mortality rates
# by arm
#-------------------------------
dcl3 %>%
  ggplot(aes(x=mrate*1000, color=tx, fill=tx))+
  geom_histogram(binwidth=1,alpha=0.5)+
  scale_color_manual(values=c("#999999", "#E69F00"))+
  scale_fill_manual(values=c("#999999", "#E69F00"))+
  scale_x_continuous(breaks=seq(0,50,by=10)) + 
  scale_y_continuous(breaks=seq(0,25,by=5)) +
  facet_grid(tx ~ .) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.y = element_blank(), 
        strip.text.y = element_text(angle=0,size=12, hjust=0)) +
  labs(x="Mortality Rate per 1,000 child-years", y = "Number of clusters",
       title = "Cluster level mortality rates", subtitle = "(bin width = 1)")

#-------------------------------
# smoothed density of mortality rates
# in the two groups
#-------------------------------
dcl3$rugpos <- ifelse(dcl3$tx=="Placebo",0,-0.002)

ggplot(data=dcl3, aes(x = mrate*1000, color=tx, fill=tx)) +
  geom_density( show.legend = TRUE, fill=NA) +
  geom_point(aes(x=mrate*1000,y=rugpos),alpha=0.3, pch=21) +
  scale_color_manual(values=c("black","#E69F00")) + 
  scale_fill_manual(values=c("black","#E69F00")) + 
  scale_x_continuous(breaks = seq(0,50,by=10)) +
  labs(title="mortality rate by group", x="mortality rate (deaths per 1000 child-years)", y="density") +
  theme_minimal() +
  theme(legend.position = c(0.8,0.8)
        )
```
# Non-parametric bootstrap

## Estimate IRR with a NP bootstrap

Resample clusters with replacement

```{r 95% CI using bootstrap}
# Set the seed for reproducibility
set.seed(2023)

# Set the number of bootstrap replicates
B <- 10000
# Create a vector to store the bootstrap results
boot_results_IRR <- numeric(B)
boot_results_IRD <- numeric(B)

# Perform the bootstrap
for (i in 1:B) {
  # Sample with replacement from the dataset
  bootstrap_sample <- dcl3[sample(nrow(dcl3), replace = TRUE), ]
  # Calculate the IRR for the bootstrap sample
  boot_results_IRR[i] <- IRR_IRD(bootstrap_sample)[1]
  boot_results_IRD[i] <- IRR_IRD(bootstrap_sample)[2]
}

# Calculate the estimated IRDIRR
est_irr <- IRR_IRD(dcl3)[1]
est_ird <- IRR_IRD(dcl3)[2]

# Calculate the confidence interval for IRDIRR
boot_ci_irr <- quantile(boot_results_IRR, c(0.025, 0.975))
boot_ci_ird <- quantile(boot_results_IRD, c(0.025, 0.975))



# Print the estimated IRR and the confidence interval
cat("\n The non-parametric estimate of the IRR and its bootstrap 95% CI is:\n",
    "IRR =", sprintf("%1.2f",est_irr),
    "(",sprintf("%1.2f",boot_ci_irr[1]),
    "to",sprintf("%1.2f",boot_ci_irr[2]),")")

cat("\n The non-parametric estimate of the IRD and its bootstrap 95% CI is:\n",
    "IRR =", sprintf("%1.2f",est_ird),
    "(",sprintf("%1.2f",boot_ci_ird[1]),
    "to",sprintf("%1.2f",boot_ci_ird[2]),")")

```

# Permutation test

Conditionally permute treatment within strata, re-estimate IRR using NB1 regression, and estimate the permutation P-value using 10,000 permutations.

```{r permutation test from NB1 regression}
#-------------------------------
# run the NB-1 reg model
#-------------------------------
# placebo arm
dcl3_p <- dcl3 %>% filter(tx == "Placebo")
mnb1_p <-do.reg(dcl3_p,b1=0,qq=1,lf.generator=mk.nbpll.full)
mort_p <- mnb1_p$estimate[1]
# 0.01004351

# azithro arm
dcl3_a <- dcl3 %>% filter(tx == "Azithromycin")
mnb1_a <-do.reg(dcl3_a,b1=0,qq=1,lf.generator=mk.nbpll.full)
mort_a <- mnb1_a$estimate[1]
# 0.00821943

observed_logirr <- log(mort_a/mort_p)
# IRR = 0.818382
# log(IRR) = -0.2004

#------------------------------------------
# permute the NB1 regression
# stratified by arm, allowing heterogeneity
#------------------------------------------
nb1_irr <- function(d) {
  # Placebo NB1
  dcl3_p <- d %>% filter(tx_permuted == "Placebo")
  mnb1_p <- do.reg(dcl3_p,b1=0,qq=1,lf.generator=mk.nbpll.full)
  # save a model summary
  mort_p <- mnb1_p$estimate[1]

  # Azithromycin NB1
  dcl3_a <- d %>% filter(tx_permuted == "Azithromycin")
  mnb1_a <- do.reg(dcl3_a,b1=0,qq=1,lf.generator=mk.nbpll.full)
  # save the log IRR and its SE
  mort_a <- mnb1_a$estimate[1]
  
  # calculate logirr
  logirr = log(mort_a/mort_p)
  logirr
}

#-------------------------------
# conditional permutation function
# permute treatment within each 
# of 2 randomization strata
#
# this is very much tailored to our
# analysis dataset, but it is simple
#
# it stratifies the data.frame, 
# re-permutes treatment within strata (hdss_strata),
# and then returns the data.frame with permuted treatment
# stored in a new variable, tx_permuted
#
# arguments:
# @x a cluster-level data frame with variables tx, hdss_strata
#    where tx is the treatment label and hdss_strata indicates randomization strata
#
# returns the same data frame, but with treatment shuffled
#-------------------------------
conditionally_permute <- function(x) {
    tx0 <- x[x$hdss_strata == 0,]
    tx1 <- x[x$hdss_strata == 1,]
    tx0$tx_permuted <- sample(tx0$tx,size = nrow(tx0), replace=FALSE)
    tx1$tx_permuted <- sample(tx1$tx,size = nrow(tx1), replace=FALSE)
    return(bind_rows(tx0,tx1))
}

#-------------------------------
# run the function N times to 
# produce the null distribution of the log IRR
#-------------------------------
# Set the number of permutations
N <- 10000

# Vector to store the permuted log incidence rate ratios
permuted_logirr <- numeric(N)

# Perform the permutation test in parallel using foreach and %dopar%
permuted_logirr <- foreach(i = 1:N, .combine = c) %dopar% {
  set.seed(i)
  # Permute tx randomly
  di <- conditionally_permute(dcl3)
  # Calculate the logirr for the permuted data
  nb1_irr(di)
}

# Calculate the p-value by comparing the observed statistic to the permutation distribution
p_value <- sum(abs(permuted_logirr) >= abs(observed_logirr)) / N

cat("Permutation P-value = ", round(p_value, 3), "\n")
```

# Summary table

```{r summary mortality table, warning = FALSE, message = FALSE}
#-------------------------------
# summary of deaths and person-time
#-------------------------------
dtable <- dcl3 %>%
  # tabulate deaths and person time by group
  group_by(tx) %>%
  summarise(ndeaths = sum(ndeaths), py = sum(person_years), .groups = "keep") %>%
  # calculate cumulative mortality rate per 1000
  mutate(mrate= (ndeaths / py)*1000) %>%
  pivot_wider(names_from=tx, values_from=c("ndeaths", "py", "mrate")) %>%
  dplyr::select(
         ndeaths_Azithromycin, py_Azithromycin, mrate_Azithromycin,
         ndeaths_Placebo, py_Placebo, mrate_Placebo
         )

# prepare the IRR from the model and p-value
df_irr <- data.frame(irr=paste0(sprintf("%1.2f",est_irr), 
                                   " (", sprintf("%1.2f",boot_ci_irr[1]), 
                                   ", ", sprintf("%1.2f",boot_ci_irr[2]), ")"),
                     ird=paste0(sprintf("%1.2f",est_ird), 
                                   " (", sprintf("%1.2f",boot_ci_ird[1]), 
                                   ", ", sprintf("%1.2f",boot_ci_ird[2]), ")"),
                        pvalue=p_value)

# bind everything together
dtable <- bind_cols(data.frame(outcome="All-cause Mortality"), 
                    dtable, df_irr)

# render table
knitr::kable(dtable, 
             digits = c(0,0,0,2,0,0,2,0,0,3),
             format.args = list(big.mark = ","),
             col.names = c("Outcome", 
                           "N Deaths", "Person-Years", "Rate per 1,000 PY", 
                           "N Deaths", "Person-Years", "Rate per 1,000 PY", 
                           "IRR (95% CI)*", "IRD (95% CI)*", "P-value†"),
             caption = "Incidence rate of all-cause mortality and 
             incidence rate ratio (IRR) for comparison between groups."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 3)) %>%
  footnote(symbol = c("95% CI estimated using a non-parametric bootstrap", 
                      "Permutation P-value (10,000 replicates) using NB1 regression model"))
```
```{r, warning= FALSE, message = FALSE}
df_irr <- data.frame(irr_est = est_irr,
                     irr_lower = boot_ci_irr[1],
                     irr_upper = boot_ci_irr[2])

plot_irr_horizontal <- df_irr %>%
 ggplot(data=., aes(y = 1, x=irr_est, xmin=irr_lower, xmax=irr_upper))+
  geom_vline(xintercept=1, linetype="dashed", size=1.5, col="grey40") +
  geom_pointrange(aes(), lwd=1, col="black") +
  scale_x_continuous(trans = log_trans(), breaks = seq(0.5,1.5,by=0.25), limits = c(0.5,1.5)) +
  labs( y="", x="IRR") +
  theme_bw() +
  theme(legend.position = "none", 
        strip.text.x = element_text(angle=0,size=12,hjust=0),
        strip.background = element_rect(fill=NA),
        axis.text = element_text(size=14),
        axis.text.y = element_blank(),
        axis.title = element_text(size=14),
        panel.border = element_blank() # Removes the box around the figures
        )

# save plot
ggsave(plot_irr_horizontal, 
       filename= paste0(here(), "/output/figures/CHAT_mortality_plot.jpeg"),
       width=3, height=2)
ggsave(plot_irr_horizontal, 
       filename= paste0(here(), "/output/figures/CHAT_mortality_plot.eps"),
       width=3, height=2)
```



# Session Info
```{r session info}
sessionInfo()
```

