---
title: "CHAT Trial Primary Analysis - Supplementary adjusted analysis"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

During peer review, referees suggested a supplementary analysis to estimate the relative reduction in mortality after adjusting for potentially prognostic child- and cluster-level characteristics. 

We complete this analysis, below. Note: this analysis was not prespecified in the trial's Statistical Analysis Plan and is sharply distinguisthed from the primary analysis, which estimates the incidence rate ratio using a non-parametric estimator.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here::here("R","CHAT-primary-Config.R"))
```

# Load the data

Load the final child-phase level dataset for the CHAT trial. 


```{r load the data}
#---------------------------------
# load the preprocessed data
#---------------------------------

# cluster-phase level data
# for getting an estimate of cluster size
dcl <- read_rds(here("data","CHAT_cluster_phase_public.rds"))

# child-phase level data
dchp <- read_rds(here("data","CHAT_child_phase_public.rds"))

# treatment allocations
tx_alloc <- read_rds(here("data","CHAT_tx_alloc_public.rds")) %>% 
  dplyr::select(clusterid_public, hdss_strata = mortplus)

# merge on the treatment allocation and strata to the main cluster-level summaries
dchp <- dchp %>%
  left_join(tx_alloc, by="clusterid_public") %>%
  left_join(dcl %>% 
              ungroup() %>% 
              select(clusterid_public, phase, natrisk), 
            join_by("clusterid_public","phase")) %>%
  # scale CSPS distance by 1000
  mutate(csps_distance = csps_distance/1000)


```


```{r process the data}
#------------------------------
# convert person-days into
# person-years for convenience
#------------------------------
dchp2 <- dchp %>%
  mutate(person_years = person_days/365.25)

#------------------------------
# for individual child-phase data:
#
# limit to child-phases where children 
# were at risk and were not lost
#------------------------------
table(dchp2$atrisk, dchp2$lost)
dchp3 <- dchp2 %>%
  filter(atrisk == 1, lost == 0) %>%
  mutate(phase_year = case_when(
    phase %in% c("0to6","6to12") ~ "0to12",
    phase %in% c("12to18","18to24") ~ "12to24",
    phase %in% c("24to30","30to36") ~ "24to36"),
    phase_year = factor(phase_year)
         )  %>%
  rename(month=monthofyear)

# MUAC is missing for a large number of children
# particularly 1-11m olds
# (n=3124 child-phases missing 1-11m, total missing is 3853 child-phases)
dchp3 %>% 
  group_by(agecat) %>% 
  summarize(mean_muac = mean(muac, na.rm=TRUE), na_muac = sum(is.na(muac)))

```


# Unadjusted Analysis

Estimate the unadjusted IRR, for an internal consistency check with the primary analysis

```{r np irr ird estimate, warning = FALSE}
#-------------------------------
# bootstrap resample clusters with replacement
#
# within each bootstrap replicate, 
# estimate the age-stratified
# Incidence Rate Ratio (IRR) and the
# Incidence Rate Difference
#
# summarize the percentile-based
# 95% CI for the IRR and IRD
#-------------------------------

# Define the bootstrap function to calculate IRR and IRD
IRR_IRD <- function(d) {
  az_person_years <- sum(d$person_years[d$tx == "Azithromycin"])
  pl_person_years <- sum(d$person_years[d$tx == "Placebo"])
  az_deaths <- sum(d$death[d$tx == "Azithromycin"])
  pl_deaths <- sum(d$death[d$tx == "Placebo"])
  az_incidence <- 1000*az_deaths / az_person_years
  pl_incidence <- 1000*pl_deaths / pl_person_years
  irr <- az_incidence / pl_incidence
  ird <- (az_incidence - pl_incidence)
  return(list(irr = irr, ird = ird, az_incidence = az_incidence, pl_incidence = pl_incidence))
}

# Set the number of bootstrap replicates (set to 10,000 for publication-quality 95% CIs)
bootreps <- 10000

# draw a boostratp sample, resampling clusters with replacement
clids <- unique(dchp3$clusterid_public)
bsamp <- matrix(sample(clids,length(clids)*bootreps, replace=TRUE),
                nrow=length(clids),ncol=bootreps )
              
# bootstrap the IRR and IRD on the child-phase level dataset
irr_ird_boot <- foreach(booti = 1:bootreps, .combine = rbind) %dopar% {
  # set seed within dopar loop for perfect reproducibility
  set.seed(booti)
  
  # draw the bootstrap sample, estimate non-parametric IRR and IRD, return results
  di <- left_join( data.frame(clusterid_public=bsamp[,booti]), dchp3, by="clusterid_public")
  results <- IRR_IRD(di)
  return(data.frame(bootrep = booti, 
                    irr=results$irr, ird=results$ird, 
                    az_incidence = results$az_incidence, 
                    pl_incidence = results$pl_incidence)
  )
  
}

# summarize estimates and their 95% CI
est_irr_ird <- IRR_IRD(dchp3)

  # Calculate the confidence interval
  boot_irr_ci <- quantile(irr_ird_boot$irr, c(0.025, 0.975))
  boot_ird_ci <- quantile(irr_ird_boot$ird, c(0.025, 0.975))
  boot_azir_ci <- quantile(irr_ird_boot$az_incidence, c(0.025, 0.975))
  boot_plir_ci <- quantile(irr_ird_boot$pl_incidence, c(0.025, 0.975))

  # Print the estimated IRR and the confidence interval
  cat("\nThe non-parametric estimate of the IRR and its bootstrap 95% CI is:\n",
      "IRR =", sprintf("%1.2f",est_irr_ird$irr),"(",sprintf("%1.2f",boot_irr_ci[1]),"to",sprintf("%1.2f",boot_irr_ci[2]),")\n",
      "IRD =", sprintf("%1.2f",est_irr_ird$ird),"(",sprintf("%1.2f",boot_ird_ci[1]),"to",sprintf("%1.2f",boot_ird_ci[2]),")")

```

# Adjusted analysis


## Estimate unadjusted IRR and robust 95% CI

Use a Poisson regression with robust SEs clustered at the cluster level to estimate the IRR and 95% CI. This is for direct comparison with the non-parametric estimator, above, to ensure consistency. 

```{r unadjusted modified poisson model}
#-------------------------------
# Unadjusted Poisson Model
# child-phase level data
#-------------------------------
irr_unadj <- glm(death ~ tx, data=dchp3, family=poisson(link="log"))
( irr_unadj_rb <- coeftest(irr_unadj, vcov.=vcovCL(irr_unadj,cluster=dchp3$clusterid_public)) )

cat("\nThe unadjusted GLM estimate of IRR and its robust 95% CI is:\n",
    "IRR =", sprintf("%1.2f",exp(irr_unadj_rb[2,1])),"(",sprintf("%1.2f", exp(irr_unadj_rb[2,1] - 1.96*irr_unadj_rb[2,2])),"to",sprintf("%1.2f", exp(irr_unadj_rb[2,1] + 1.96*irr_unadj_rb[2,2])),")\n"
    )
```

## Estimate adjusted IRR and robust 95% CI

Use a Poisson regression model to complete an adjusted analysis, adding child- and cluster-level characteristics.

```{r adjusted estimate }
#-------------------------------
# Adjusted Poisson model
#-------------------------------
irr_adj <- glm(death ~ tx + age_months + gender + csps_distance + natrisk + month, data=dchp3, family=poisson(link="log"))

cat("\n Adjusted model fit (SEs not adjusted for clustering!)\n")
summary(irr_adj)

cat("\n Adjusted model fit, cluster robust SEs\n")
( irr_adj_rb <- coeftest(irr_adj, vcov.=vcovCL(irr_adj,cluster=dchp3$clusterid_public)) ) 

# print the exponentiated coefficients (IRR) along with the robust 95% CIs
irr_adj_tab <- cbind(irr=exp(irr_adj_rb[,1]),exp(confint(irr_adj_rb)))[-1,]
kbl(irr_adj_tab, caption = "Adjusted model estimate of the IRR and robust 95% CIs", digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")

cat("\nThe adjusted GLM estimate of IRR and its robust 95% CI is:\n",
    "IRR =", sprintf("%1.2f",exp(irr_adj_rb[2,1])),"(",sprintf("%1.2f", exp(irr_adj_rb[2,1] - 1.96*irr_adj_rb[2,2])),"to",sprintf("%1.2f", exp(irr_adj_rb[2,1] + 1.96*irr_adj_rb[2,2])),")\n"
    )
```

Additionally add MUAC, despite missing values for 3853 child-phases

```{r adjusted estimate also including MUAC}
#-------------------------------
# Adjusted Poisson model
# adjusting for MUAC, month of year, and distance to CSPS
# note: drops many observations due to missing covariates
#-------------------------------
irr_adj2 <- glm(death ~ tx + age_months + gender + csps_distance + natrisk + muac + month, data=dchp3, family=poisson(link="log"))

cat("\n Adjusted model fit (SEs not adjusted for clustering!)\n")
summary(irr_adj2)

cat("\n Adjusted model fit, cluster robust SEs\n")
( irr_adj2_rb <- coeftest(irr_adj2, vcov.=vcovCL(irr_adj2,cluster=dchp3$clusterid_public)) )


# print the exponentiated coefficients (IRR) along with the robust 95% CIs
irr_adj2_tab <- cbind(irr=exp(irr_adj2_rb[,1]),exp(confint(irr_adj2_rb)))[-1,]
kbl(irr_adj2_tab, caption = "Adjusted model estimate of the IRR and robust 95% CIs", digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")


cat("\nThe adjusted GLM estimate of IRR and its robust 95% CI is:\n",
    "IRR =", sprintf("%1.2f",exp(irr_adj2_rb[2,1])),"(",sprintf("%1.2f", exp(irr_adj2_rb[2,1] - 1.96*irr_adj2_rb[2,2])),"to",sprintf("%1.2f", exp(irr_adj2_rb[2,1] + 1.96*irr_adj2_rb[2,2])),")\n"
    )
```
## Permutation P-value for the adjusted IRR

```{r permutation test of the adjusted model}

#-------------------------------
# conditional permutation function
# permute treatment within each 
# of 2 randomization strata
#
# this is very much tailored to our
# analysis dataset, but it is simple
#
# it stratifies the data.frame, 
# re-permutes treatment within strata (hdss_strata),
# and then returns the data.frame with permuted treatment
# stored in a new variable, tx_permuted
#
# arguments:
# @x a cluster-level data frame with variables tx, hdss_strata
#    where tx is the treatment label and hdss_strata indicates randomization strata
#
# returns the same data frame, but with treatment shuffled
#-------------------------------
conditionally_permute <- function(x) {
    tx0 <- x[x$hdss_strata == 0,]
    tx1 <- x[x$hdss_strata == 1,]
    tx0$tx_permuted <- sample(tx0$tx,size = nrow(tx0), replace=FALSE)
    tx1$tx_permuted <- sample(tx1$tx,size = nrow(tx1), replace=FALSE)
    return(bind_rows(tx0,tx1))
}


#-------------------------------
# conduct a permutation test
# of the adjusted log(IRR)
# estimated within the Poisson model
# of child-phase level data
# permute treatement at the 
# cluster level
#-------------------------------

# Set the number of permutations to use (set to 10,000 for publication-quality P-value)
Npermute <- 10000

adj_irr_permutations <- foreach(permi = 1:Npermute, .combine = rbind) %dopar% {
  # Set the seed for reproducibility
  set.seed(permi)
  
  # Permute the treatment at the cluster level
  dtx <- dchp3 %>% select(clusterid_public, tx, hdss_strata) %>% group_by(clusterid_public) %>% slice(1) %>% ungroup() %>% conditionally_permute()
  
  di <- left_join(dtx, dchp3, by=c("clusterid_public","tx","hdss_strata"))
  
  # Fit the adjusted Poisson model
  irr_adj <- glm(death ~ tx_permuted + age_months + gender + csps_distance + natrisk + month, data = di, family = poisson(link = "log"))
  # return the adjusted log(IRR), which is the second coeff (after the intercept)
  return(irr_adj$coefficients[2])
}

  
# Calculate the permutation P-value
pval_perm_adj_irr <- sum(abs(adj_irr_permutations) >= abs(irr_adj_rb[2,1])) / length(adj_irr_permutations)

# Print the results
cat("Permutation P-value:", pval_perm_adj_irr, "\n")
```

Additional permutation test, also including MUAC in the adjustment set


```{r permutation test of the adjusted model also including MUAC}

#-------------------------------
# conduct a permutation test
# of the adjusted log(IRR)
# estimated within the Poisson model
# of child-phase level data
# permute treatement at the 
# cluster level
#-------------------------------

# Set the number of permutations to use (set to 10,000 for publication-quality P-value)
Npermute <- 10000

adj_irr_permutations <- foreach(permi = 1:Npermute, .combine = rbind) %dopar% {
  # Set the seed for reproducibility
  set.seed(permi)
  
  # Permute the treatment at the cluster level
  dtx <- dchp3 %>% select(clusterid_public, tx, hdss_strata) %>% group_by(clusterid_public) %>% slice(1) %>% ungroup() %>% conditionally_permute()
  
  di <- left_join(dtx, dchp3, by=c("clusterid_public","tx","hdss_strata"))
  
  # Fit the adjusted Poisson model
  irr_adj <- glm(death ~ tx_permuted + age_months + gender + csps_distance + natrisk + muac + month, data = di, family = poisson(link = "log"))
  # return the adjusted log(IRR), which is the second coeff (after the intercept)
  return(irr_adj$coefficients[2])
}

  
# Calculate the permutation P-value
pval_perm_adj_irr <- sum(abs(adj_irr_permutations) >= abs(irr_adj_rb[2,1])) / length(adj_irr_permutations)

# Print the results
cat("Permutation P-value:", pval_perm_adj_irr, "\n")
```

# Session info
```{r session info}
sessionInfo()
```

